Wealth Inequality in China: Evidence from the 2017 and 2019 CHFS

### 2025-04-23 report/draft.qmd
update

The factor decompostion of the Gini coefficient for household wealth in two years shows that housing equity is still the largest and dominant contributor to China's wealth inequality, contributing around 70% to the overall wealth Gini. Financial equity follows by contributing to 11.1% in 2017 and 12.5% in 2019 to the overall wealth Gini. Commercial equity such as agricultural and business assets contrbutes to a relatively large portion in 2017 but to a smaller portion in 2019. Land asset contributes to a relatively large portion in 2017 but to a smaller portion in 2019. The cause of this is uninvestigated, which could be due to the change in survey design. Commercial property and car equity contribute to a relatively small portion in both years. Educational debt and medical debt negatively correlates with wealth inequality, which shows that the poor is disporportionately burdened by educational and medical debts, although their contribution to the overall wealth Gini is small.

This high contribution of housing equity to wealth inequality is consistent with other studies. @xie_household_2015 used the 2012 CHPS data and documented that housing equity conprises of 71.6% of household wealth in China. @kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013. @tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the same ratio was 71.08% in 2017 and 67.23% in 2019. It appears that housing remains the largest wealth holding for Chinese households in 2017 and 2019, and housing is still the largest contributor to wealth inequality with no clear sign of change.

>>>>

The factor decompostion of the Gini coefficient for household wealth in two years shows that housing asset dominates the wealth inequality in China. In both 2017 and 2019, the housing asset occupy over 70% of total household wealth and contribute to over 72% of the overall wealth Gini. This high contribution of housing equity to wealth inequality is consistent with other studies. @tan_earnings_2017 documented with the 2011 CHFS data that the housing asset to total asset ratio was 66.73%, and the 2017 and 2019 74.2% and 73% shows an sharp increase in the housing asset to total asset ratio from 2011 to 2019. Meanwhile, @kanbur_great_2021 used the 2002 and 2013 CHIP data and found that housing euqity contribues to 64% of wealth inequality in 2002 and 79% in 2013. Albeit variation in the statistics across sources, the results in @tbl-tbl3 demonstrates that in 2017 and 2019, the housing market is still the most important driver of wealth inequality in China. However, surprisingly,
